---
title: "Exeter Density"
author: "Bryan Ellerbrock"
date: "11/2/2021"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/projects/RScripts/Blockiness")
cdata <- read.csv("Exeter_calibration.csv", header = TRUE, check.names = TRUE,  stringsAsFactors = FALSE)
```

## Calibration Data

```{r echo=FALSE}
knitr::kable(cdata, format="markdown")
```

length in inches x 100 and weight in ounces x 10

## Calculating Density

Multiplying length by 0.0254 to get lengthCm (0.0254cm per 100th of an inch)

```{r echo=FALSE}
cdata$lengthCm = cdata$length * 0.0254 # 0.0254 cm per 100th of an inch
knitr::kable(cdata[,c(1,5)], format="markdown")
```

Dividing pixelLength by lengthCm to get pixels/cm estimates

```{r echo=FALSE}
cdata$pixelsPerCm = cdata$pixelLength /cdata$lengthCm
knitr::kable(cdata[,c(2,5,6)], format="markdown")
```

Calculating mean number of pixels/cm, then raising to power of three to get cubic pixels/cubic cm

```{r}
meanPixelsPerCm = mean(cdata$pixelsPerCm)
cubedPixelsPerCmCubed = meanPixelsPerCm ^ 3
```

Dividing pixelVolume by cubic pixels/cubic cm to get volume in cubic cm

```{r echo=FALSE}
cdata$cmCubed = cdata$pixelVolume / cubedPixelsPerCmCubed
knitr::kable(cdata[,c(4,7)], format="markdown")
```

Multiplying weight by 2.83495 to get weight in grams (2.83495 g per 10th of an oz)

```{r echo=FALSE}
cdata$weightGrams = cdata$weight * 2.83495
knitr::kable(cdata[,c(3,8)], format="markdown")
```

Dividing weight in grams by volume in cubic cms to get density estimates, then calculating mean density

```{r echo=FALSE}
cdata$densityCmCubed = cdata$weightGrams / cdata$cmCubed
knitr::kable(cdata[,c(7,8,9)], format="markdown")
```

```{r}
meanDensityCmCubed= mean(cdata$densityCmCubed)
meanDensityCmCubed
```
